---
title: "p8105_hw1_mp3745"
author: "Matthew Parker"
date: "9/13/2019"
output: github_document
---

# Problem 1

```{r, message=FALSE}
# Load tidyverse library
library(tidyverse)
```


```{r}
# Create first dataframe
prob1_df <- tibble(
  norm_samp = rnorm(8, mean = 8),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h"),
  vec_factor = factor(c("burrito", "quesadilla", "tacos", "burrito", "burrito", "tacos", "tacos", "tacos"))
)

# See if mean can be calculated for norm_samp in prob1_df
mean(pull(prob1_df, norm_samp))

# See if mean can be calculated for norm_samp_pos in prob1_df
mean(pull(prob1_df, norm_samp_pos))

# See if mean can be calculated for vec_char in prob1_df
mean(pull(prob1_df, vec_char))

# See if mean can be calculated for vec_factor in prob1_df
mean(pull(prob1_df, vec_factor))
```

The mean could be calculated for norm_samp and norm_samp_pos, but not vec_char or vec_factor in prob1_df.



```{r, eval=FALSE}
# Apply as.numeric to norm_samp_pos from prob1_df
as.numeric(pull(prob1_df, norm_samp_pos))

# Apply as.numeric to vec_char from prob1_df
as.numeric(pull(prob1_df, vec_char))

# Apply as.numeric to vec_factor from prob1_df
as.numeric(pull(prob1_df, vec_factor))
```

In the logical vector, all the "TRUE"s were converted into 1's. In both the character and factor vector, everything was converted to "NA". This explains how we were able to calculate a mean for the logical vector, but not the character or factor vector.



```{r}
# Convert the logical vector to numeric, and multiply the random sample by the result
as.numeric(pull(prob1_df, norm_samp_pos)) * pull(prob1_df, norm_samp)

# convert the logical vector to a factor, and multiply the random sample by the result
factor(pull(prob1_df, norm_samp_pos)) * pull(prob1_df, norm_samp)

# convert the logical vector to a factor and then convert the result to numeric, and multiply the 
# random sample by the result
as.numeric(factor(pull(prob1_df, norm_samp_pos))) * pull(prob1_df, norm_samp)
```



# Problem 2

```{r}
# Create the dataframe
prob2_df <- tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_log = x + y > 1,
  vec_num = as.numeric(vec_log),
  vec_fac = factor(vec_log)
)
```

The above dataframe (prob2_df) has `r nrow(prob2_df)` rows and `r ncol(prob2_df) ` columns. The mean of x is `r mean(pull(prob2_df, x))`. The median of x is `r median(pull(prob2_df, x))`. The standard deviation of x is `r sd(pull(prob2_df, x))`. In addition, the proportion of cases in which x + y > 1 is `r sum(pull(prob2_df, vec_log)) / length(pull(prob2_df, vec_log))`.

```{r}
# Scatterplot of y vs x by vec_log from prob2_df
scat_plot_log <- ggplot(prob2_df, aes(x = x, y = y, color = vec_log)) +
  geom_point()

# Scatterplot of y vs x by vec_num from prob2_df
scat_plot_num <- ggplot(prob2_df, aes(x = x, y = y, color = vec_num)) +
  geom_point()

# The color scale in the above plot is a sliding scale from blue (vec_num = 1) to black (vec_num = 0), 
# but there are only two distinct colors because vec_num only has two values (0 and 1).

# Scatterplot of y vs x by vec_fac from prob2_df
scat_plot_fac <- ggplot(prob2_df, aes(x = x, y = y, color = vec_fac)) +
  geom_point()

# The color scale in the above plot is the same as when coloring by the logical vector. The factor 
# vector only has two values (True or False).

# Export the first scatterplot
ggsave("scatter_plot_log.png", plot = scat_plot_log)

```

